---
title: "Logit Model and Accompanying Figures"
date: "3/5/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, eval = T, error = T, warning = FALSE, message=FALSE)
suppressMessages(library(ggplot2))
suppressMessages(library(tidyverse))
suppressMessages(library(dplyr))
```

```{r}
setwd("~/Desktop/Junior/19W/GOVT19/GOVT FINAL/Final_Gov19_Submission")
#load(file = "~/Desktop/GOV19/Final Project/inmates.Rdata")
detach("package:dplyr", character.only = TRUE)
library("dplyr", character.only = TRUE)
load("inmates.Rdata")
inmates_logit <- inmates %>% select(AGE_CAT, RACE, SES_IMEDFAMILY_INCARCERATED, EDUCATION, GENDER, 
                                    SENTENCEMTH, LIFE_DEATH, OFFENSE)
```

#Recoding variables
```{r}
#Create binary female variable 
inmates_logit$FEMALE <- NA
inmates_logit$FEMALE[inmates_logit$GENDER == "Male"] <- 0
inmates_logit$FEMALE[inmates_logit$GENDER == "Female"] <- 1

#Create four dummy variables for each race 
inmates_logit$RACE[inmates_logit$RACE == "Missing"] <- NA
inmates_logit$BLACK <- NA
inmates_logit$HISPANIC <- NA 
inmates_logit$WHITE <- NA 
inmates_logit$OTHER <- NA 
inmates_logit$BLACK[inmates_logit$RACE != "Black"] <- 0 
inmates_logit$BLACK[inmates_logit$RACE == "Black"] <- 1
inmates_logit$WHITE[inmates_logit$RACE != "White"] <- 0 
inmates_logit$WHITE[inmates_logit$RACE == "White"] <- 1
inmates_logit$HISPANIC[inmates_logit$RACE != "Hispanic"] <- 0 
inmates_logit$HISPANIC[inmates_logit$RACE == "Hispanic"] <- 1
inmates_logit$OTHER[inmates_logit$RACE != "Other"] <- 0 
inmates_logit$OTHER[inmates_logit$RACE == "Other"] <- 1

#Recode education so that "Missing" values become NA's that the model then removes
inmates_logit$EDUCATION[as.numeric(inmates_logit$EDUCATION) == 19] <- NA

#Recode the factor variable of education to a new numeric variable
inmates_logit <- inmates_logit %>% mutate(EDUCATION_NUM = as.numeric(EDUCATION))

#Recode socioeconomic status variable
inmates_logit$FAMILY_INCARCERATED <- NA
inmates_logit$FAMILY_INCARCERATED[inmates_logit$SES_IMEDFAMILY_INCARCERATED == "Immediate family never incarcerated"] <- 0 
inmates_logit$FAMILY_INCARCERATED[inmates_logit$SES_IMEDFAMILY_INCARCERATED == "Immediate family incarcerated"] <- 1 
inmates_logit$FAMILY_INCARCERATED[inmates_logit$SES_IMEDFAMILY_INCARCERATED == "Missing"] <- NA

#Recode age category variable into a 5 point scale variable
inmates_logit$AGE <- NA
inmates_logit$AGE[inmates_logit$AGE_CAT == "< 25 yrs"] <- 0
inmates_logit$AGE[inmates_logit$AGE_CAT == "25-34"] <- 1
inmates_logit$AGE[inmates_logit$AGE_CAT == "35-44"] <- 2
inmates_logit$AGE[inmates_logit$AGE_CAT == "45-54"] <- 3
inmates_logit$AGE[inmates_logit$AGE_CAT == "55-64"] <- 4
inmates_logit$AGE[inmates_logit$AGE_CAT == "65-96"] <- 5

#Remove observations for which we have no sentencing data at all.  
#These are cases where SENTENCEMTH is NA and LIFE_DEATH is FALSE.
#To do this, we make a 'missing sentence' column, where 1 will represent entries that have no sentencing data.
inmates_logit$MISSINGSENTENCE <- NA
inmates_logit$MISSINGSENTENCE[is.na(inmates_logit$SENTENCEMTH) & inmates_logit$LIFE_DEATH == FALSE] <- 1
inmates_logit$MISSINGSENTENCE[inmates_logit$SENTENCEMTH >= 0] <- 0
inmates_logit$LIFE_DEATH2 <- NA
inmates_logit$LIFE_DEATH2[inmates_logit$LIFE_DEATH == FALSE] <- 0 
inmates_logit$LIFE_DEATH2[inmates_logit$LIFE_DEATH == TRUE] <- 1
inmates_logit <- inmates_logit %>% 
  select(LIFE_DEATH2, SENTENCEMTH, LIFE_DEATH, everything()) %>% 
  filter(MISSINGSENTENCE != 1 | is.na(MISSINGSENTENCE))
```

#Subset data to only include entries for violent offenses, as violent offenses are a near perfect predictor for receiving extreme sentences.
```{r}
inmates_logit <- subset(inmates_logit, OFFENSE == "VIOLENT")
```

#Create a logit model for the binary variable of receiving an extreme sentence (sentenced to life or death) or not.
```{r}
logit.model <- glm(LIFE_DEATH2 ~ BLACK + HISPANIC + OTHER + FEMALE + EDUCATION_NUM + FAMILY_INCARCERATED + AGE,
              data = inmates_logit, family = binomial(link = "logit"))
summary(logit.model)
exp(coef(logit.model)) / (1 + exp( coef(logit.model) ))
```

# Calculate percent correctly predicted by the model.
```{r}
library(pscl)
hitmiss(logit.model)
```

#Create a stargazer table of the logistic regression model.
```{r}
library(stargazer)
stargazer(logit.model, type = "text", title = "Change in Pr(Receiving an Extreme Sentence) based on Predictors",
          covariate.labels = c("Black", "Hispanic","Other Race", "Female", "Education","Immediate Family Incarcerated","Age"),
          dep.var.labels = c("Change in Pr(Extreme Sentence)"),
          add.lines = list(c("Percent correctly predicted", "84.69%")),
          star.char = "*",
          star.cutoffs = 0.05, 
          notes = c("*p < .05, two-tailed"),
          notes.append = FALSE,
          header = FALSE)
```

#Make an airplane plot of the change in predicted probability for a unit change in predictors.
```{r}
library(gtools)
#library(lme4)
#library(arm)

# Calculate predicted probability for a profile
# A white male HS graduate with no immediate family incarcerated and under 25 years old
pred_logit <- logit.model$coef[1] + logit.model$coef[2] * 0 +
  logit.model$coef[3] * 0 + logit.model$coef[4] * 0 + logit.model$coef[5] * 0 +
  logit.model$coef[6] * 13 + logit.model$coef[7] * 0 + logit.model$coef[8] * 0
inv.logit(pred_logit)

# Calculate that same profile except race changed to black (to check to see if airplane plot is plotting correctly)
pred_logit_black <- logit.model$coef[1] + logit.model$coef[2] * 1 +
  logit.model$coef[3] * 0 + logit.model$coef[4] * 0 + logit.model$coef[5] * 0 +
  logit.model$coef[6] * 13 + logit.model$coef[7] * 0 + logit.model$coef[8] * 0
inv.logit(pred_logit_black)

```


```{r}
covariates <- c("Black" ,"Hispanic", "Other", "Female", "Education", "Immediate Family\nIncarcerated", "Age")
logit.model$coef[1]
logit_beta <- c(logit.model$coef[2],
                logit.model$coef[3],
                logit.model$coef[4],
                logit.model$coef[5],
                logit.model$coef[6],
                logit.model$coef[7],
                logit.model$coef[8])

logit_se <- c(sqrt(vcov(logit.model)[2,2]),
              sqrt(vcov(logit.model)[3,3]),
              sqrt(vcov(logit.model)[4,4]),
              sqrt(vcov(logit.model)[5,5]),
              sqrt(vcov(logit.model)[6,6]),
              sqrt(vcov(logit.model)[7,7]),
              sqrt(vcov(logit.model)[8,8]))

logit_df <- data.frame(logit_beta, logit_se)
logit_df

logit_df$fd <- (inv.logit(pred_logit + logit_beta) - inv.logit(pred_logit))
logit_df$fdup <- (inv.logit(pred_logit + logit_beta + 1.96 * logit_se) - inv.logit(pred_logit))
logit_df$fdlb <- (inv.logit(pred_logit + logit_beta - 1.96 * logit_se) - inv.logit(pred_logit))

ggplot(data = logit_df, 
       aes(x = reorder(covariates, fd), y = fd)) + 
  geom_pointrange(aes(ymin = fdlb, ymax = fdup), width = .1) + coord_flip() +
  ggtitle("Effects Plot of Change in Probability of Receiving an Extreme Sentence") + 
  labs(y = "Change in Probability of Receiving an Extreme Sentence\nDue to One Unit Change in Predictor",
       x = "Predictor") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_y_continuous(breaks = seq(-0.02, 0.04, 0.01), limits = c(-0.02, 0.04))
```

# Make a graph of the probability of receiving an extreme sentence as a function of age.
```{r}
par(mar=c(5,6,4,1)+.1)
sim.data <- data.frame(BLACK = 0,
                       HISPANIC = 0,
                       OTHER = 0,
                       FEMALE = 0,
                       EDUCATION_NUM = 13,
                       FAMILY_INCARCERATED = 0,
                       AGE = c(1,2,3,4,5,6))
summary(sim.data)

OutHats <- predict(logit.model, se.fit = TRUE, newdata = sim.data)
OutHats
OutHatsUB <- OutHats$fit + (1.96 * OutHats$se.fit)
OutHatsLB <- OutHats$fit - (1.96 * OutHats$se.fit)
OutHats <- cbind(as.data.frame(OutHats), OutHatsUB, OutHatsLB)
OutHats <- data.frame(lapply(OutHats, binomial(link = "logit")$linkinv))
both <- cbind(sim.data, OutHats)
```

```{r}
plot(both$AGE,
     both$fit,
     ylim = c(0, 1),
     t = "l",
     lwd = 2,
     xlab = "Age",
     ylab = "",
     main = "Pr(Extreme Sentence) as a Function of Age"
     ,axes = F
     )
lines(both$AGE, both$OutHatsUB, lty = 2)
lines(both$AGE, both$OutHatsLB, lty = 2)
axis(side = 1, labels = NA)
axis(side = 2)
lablist = c("\n< 25", "\n25-34", "\n35-44", "\n45-54", "\n55-64", "\n65-96")
text(seq(1, 6, by = 1), par("usr")[3], labels = lablist, srt = 0, pos = 1, xpd = T, cex = 1)
title(ylab = "Predicted Probability of Defendant\nReceiving an Extreme Sentence", line = 2.3, cex.lab = 1)
```

# Make a table to show the change in predicted probability of receiving an extreme sentence as the age category for the profile increases.
```{r}
library(knitr)
library(kableExtra)
library(gtools)

logit.model$coef
pred_logit <- logit.model$coef[1] + logit.model$coef[2] * 0 +
  logit.model$coef[3] * 0 + logit.model$coef[4] * 0 + logit.model$coef[5] * 0 +
  logit.model$coef[6] * 13 + logit.model$coef[7] * 0 + logit.model$coef[8] * 0
inv.logit(pred_logit)

pred_logit_2534 <- logit.model$coef[1] + logit.model$coef[2] * 1 +
  logit.model$coef[3] * 0 + logit.model$coef[4] * 0 + logit.model$coef[5] * 0 +
  logit.model$coef[6] * 13 + logit.model$coef[7] * 0 + logit.model$coef[8] * 1
inv.logit(pred_logit_2534)

pred_logit_3544 <- logit.model$coef[1] + logit.model$coef[2] * 1 +
  logit.model$coef[3] * 0 + logit.model$coef[4] * 0 + logit.model$coef[5] * 0 +
  logit.model$coef[6] * 13 + logit.model$coef[7] * 0 + logit.model$coef[8] * 2
inv.logit(pred_logit_3544)

pred_logit_4554 <- logit.model$coef[1] + logit.model$coef[2] * 1 +
  logit.model$coef[3] * 0 + logit.model$coef[4] * 0 + logit.model$coef[5] * 0 +
  logit.model$coef[6] * 13 + logit.model$coef[7] * 0 + logit.model$coef[8] * 3
inv.logit(pred_logit_4554)

pred_logit_5564 <- logit.model$coef[1] + logit.model$coef[2] * 1 +
  logit.model$coef[3] * 0 + logit.model$coef[4] * 0 + logit.model$coef[5] * 0 +
  logit.model$coef[6] * 13 + logit.model$coef[7] * 0 + logit.model$coef[8] * 4
inv.logit(pred_logit_5564)

pred_logit_6596 <- logit.model$coef[1] + logit.model$coef[2] * 1 +
  logit.model$coef[3] * 0 + logit.model$coef[4] * 0 + logit.model$coef[5] * 0 +
  logit.model$coef[6] * 13 + logit.model$coef[7] * 0 + logit.model$coef[8] * 5
inv.logit(pred_logit_6596)

inv.logit(pred_logit_2534) - inv.logit(pred_logit)
inv.logit(pred_logit_3544) - inv.logit(pred_logit_2534)
inv.logit(pred_logit_4554) - inv.logit(pred_logit_3544)
inv.logit(pred_logit_5564) - inv.logit(pred_logit_4554)
inv.logit(pred_logit_6596) - inv.logit(pred_logit_5564)

change_in_prob <- c(0.05982, 0.05398, 0.070760, 0.08735, 0.10018)
category_change <- c("Increase to 25-34 from < 25",
                     "Increase to 35-44 from 25-34",
                     "Increase to 45-54 from 35-44",
                     "Increase to 55-64 from 45-54",
                     "Increase to 65-96 from 55-64")
change_over_age <- data.frame(cbind(category_change, change_in_prob))

kable(change_over_age, format = "pandoc", booktabs = T, 
      caption = "Change in Pr(Extreme Sentence) for Age Increase Across Categories",
      escape = FALSE, col.names = c("Change in Age", "Change in Pr(Extreme Sentence)"))
```
